---
title: "Total Error: Using ML to Measure Total Exposure"
author: "Gaurav Sood"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
source("sim_total_error.R")
simulation <- simulate_exposure()
write_results(simulation)
results <- simulation$results
```

Small systematic errors can distort totals when people contribute different
numbers of observations. This notebook retains the June 2023 simulation's
seed, random draws, and original bias specification, and adds two comparisons
using the same draws. The underlying argument appears in the June 2020
[Domain Knowledge manuscript](https://github.com/themains/domain_knowledge/blob/850644c06d19a3d787a444443472367dd1bd3046/ms/domain_knowledge.tex#L180-L200).

## Design

For each of 1,000 respondents, draw an observation count uniformly from the
integers 5 through 1,000. Draw independent bivariate normal pairs with zero
means, unit variances, and correlation 0.9. Let the first component be the
true measure and the second an unbiased proxy. Compare three proxies:

- No added bias.
- A constant addition of 0.1 to every proxy observation.
- The original addition of 0.1 times the respondent's sample standard deviation
  of the true measure.

The [R source](sim_total_error.R) generates the draws and summaries. These are
continuous, centered measurements, not binary labels, visit counts, or durations.
Each respondent has independent draws; the simulation does not reproduce the
shared domain labels in browsing data. It isolates accumulation of bias.

## Results

```{r results}
display <- results
display$model <- c("No added bias", "Constant bias", "Original SD-scaled bias")
knitr::kable(display, digits = 3, col.names = c(
  "Proxy", "Observation correlation", "Mean correlation",
  "Total correlation", "Mean total error"
))
```

In the original specification, the observation-level correlation is
`r sprintf("%.3f", results$observation_correlation[3])` and the correlation of
respondent totals is `r sprintf("%.3f", results$total_correlation[3])`.
The mean signed error in totals (proxy minus truth) is
`r sprintf("%.3f", results$mean_total_error[3])`.
These are descriptive results from one seeded simulation, not estimates from
observed browsing data.

Adding the same constant to every observation leaves the correlations of
observations and respondent means exactly unchanged. Totals acquire a shift
proportional to the number of observations, which varies across respondents.
The original SD-scaled addition is only approximately constant, so its
mean correlation need not be exactly unchanged.

## Why the correlation falls in this design

Let $N$ be independent of the zero-mean pairs $(Y, Z)$, each with unit
variance and correlation $\rho$. Write $S=\sum_{r=1}^N Y_r$ and
$\widehat S=\sum_{r=1}^N(Z_r+b)$. Then

$$
\operatorname{Corr}(S,\widehat S)
=\frac{\rho}{\sqrt{1+b^2\operatorname{Var}(N)/E[N]}}.
$$

For the constant-bias design, the population correlation is
`r sprintf("%.3f", total_correlation(0.9, 0.1, 5:1000))`, compared with
`r sprintf("%.3f", results$total_correlation[2])` in this finite simulation.
If everyone contributes the same number of observations, the added bias
does not change this correlation, although it still biases every total.
The formula depends on the zero means and independence assumptions; it is
not a universal description of exposure totals.

## Correction and its limits

When the constant bias is known, subtracting $bN_i$ removes the added bias
exactly. The original specification instead requires $0.1N_i s_i$, where
$s_i$ is the respondent's sample standard deviation of the true measure.
That is an oracle correction: the simulation supplies truth that would
usually be unavailable in an application. Neither correction removes the
remaining random prediction errors.

For binary domain labels, a different correction applies. Under common
false-positive and false-negative probabilities $\alpha$ and $\beta$ that
remain valid conditional on browsing weights, predicted exposure satisfies

$$
E[\widehat T_i\mid C,y]
=\alpha V_i+(1-\alpha-\beta)T_i.
$$

Here $V_i$ is all visits and $T_i$ is visits to truly positive domains.
Thus $(\widehat T_i-\alpha V_i)/(1-\alpha-\beta)$ is unbiased when the
rates are known and their sum is not one. Estimating the rates adds uncertainty;
transferring unweighted validation rates to browsing-weighted totals requires
evidence. See the [note](total_error.pdf) for definitions, shared domain errors,
and implications for group differences.

## Reproduction

Run `make check` to test the algebra, regenerate this document and the paper's
numerical inputs, and compile the paper. The finite-enumeration tests check
the binary correction and covariance without relying on Monte Carlo tolerances.

```{r session, results='asis'}
session <- trimws(capture.output(sessionInfo()), which = "right")
cat("```text\n", paste(session, collapse = "\n"), "\n```\n", sep = "")
```
